Multi-Agent Path Finding with Matching using Increasing Cost Tree Search

### 2021-05-31 docs/paper.tex
Worked on logic for testing different solver configs systematically, worked on paper

\title{Multi-Agent Path Finding with matching using Increasing Cost Tree Search}

\author{Thom van der Woude}
\date{\today}

@@

\section{Introduction}The Dutch Railways (NS) company\footnote{\url{https://www.ns.nl/}} needs to clean and service their fleet of trains during the night in shunting yards so that they can properly bring passengers from A to B during the day. The problem of scheduling the trains to achieve this is called the Train Unit Shunting and Servicing (TUSS) problem \cite{mulderij2020}. It is an NP-hard problem with many different subproblems in addition to the basic routing, such as the scheduling of personnel and coupling and decoupling of train units entering and leaving the yard. The company is interested in learning tighter upper-bounds for the capacity of their shunting yards but a problem is that in practical scenarios, only heuristics can solve TUSS instances (as was shown by Geiger et al. \cite{geiger2018}), which gives little information about what optimal solutions would look like and what could still be achieved using existing shunting infrastructure. To gain more insight into the capacity of shunting yards and the feasibility of problem instances, Mulderij et al. \cite{mulderij2020} propose the multi-agent pathfinding (MAPF) problem \cite{stern2019} extended with a matching subproblem (hereafter MAPFM) as a suitable relaxation that can be used to compute upper-bounds for TUSS instances by solving the corresponding MAPFM instances optimally.

Comparatively, little research has been done on MAPFM, and the Conflict-Based Min-Cost-Flow algorithm by Ma and Koenig \cite{ma2016} is one of the few existing approaches to optimally solve this more general problem along with the somewhat similar one described in \cite{henkel2019}. Ma and Koenig's method builds upon the conflict-based search method and in particular on the Meta-Agent variation thereof also discussed in \cite{sharon2015} as a high-level search framework, while exploiting the connection between MAPF problems and max-flow problems (as discussed in \cite{yu2013}) in the low-level search. This begs the question: can alternative MAPF algorithms such as Increasing Cost Tree Search (ICTS) \cite{sharon2011} or M* \cite{wagner2011} also serve as the basis for a MAPFM solver?

This work is the outcome of a search for an efficient algorithm for optimally solving MAPFM derived from ICTS as an alternative to the approach outlined above\footnote{To the knowledge of the author, to date, no such algorithm is described in the literature}. ICTS, in short, is a two-level approach to MAPF with a top-level breadth-first traversal of an Increasing Cost Tree (ICT) representing combinations of per-agent path-lengths, and a bottom-level evaluation of ICT nodes using multi-value decision diagrams (MDDS) that represent per agent all possible paths to the goal of a set length. In the context of a research project in which multiple such MAPF algorithms were taken as starting points for MAPFM algorithms within a peer group, the novel ICTS-based algorithm is compared to these algorithms on a set of standard benchmarks described in \cite{stern2019}, in addition to being analysed in terms of complexity and compared to both an enumerative approach that solves all matchings using regular ICTS and the algorithm described in \cite{ma2016}. Lastly, completeness of the algorithm is considered as this is an important property in the context of MAPFM as a TUSS relaxation. Regarding the relevance of this work, it should be noted that besides TUSS, there are other applications such as planning warehouse robots \cite{wurman2007} that also could benefit from novel algorithms for MAPFM.
\printbibliography

>>>>

	\title{Multi-Agent Path Finding with Matching using Increasing Cost Tree Search}

	\author{Thom van der Woude\and Jesse Mulderij\and Mathijs de Weerdt}
	\date{}

@@

	\begin{abstract}
		Both the assignment problem and the multi-agent pathfinding problem are common problems in fields from robotics to transportation. The joint problem of finding matchings inducing an optimal non-conflicting routing of agents to their goals is something that has not been studied much. Few methods exist today that solve it although the problems do appear together in real-world problems, such as in the Train Unit Shunting and Servicing problem. In this work, a novel method based on the Increasing Cost Tree Search algorithm for multi-agent pathfinding is presented that can optimally solve this joint problem.
	\end{abstract}

	\section{Introduction}
	The Dutch Railways (NS) company\footnote{\url{https://www.ns.nl/}} needs to clean and service their fleet of trains during the night in shunting yards so that they can properly bring passengers from A to B during the day. The problem of scheduling the trains to achieve this is called the Train Unit Shunting and Servicing (TUSS) problem \cite{mulderij2020}. It is an NP-hard problem with many different subproblems in addition to the basic routing, such as the scheduling of personnel and coupling and decoupling of train units entering and leaving the yard. % introduce heuristics here.

	The company is interested in learning tighter upper-bounds for the capacity of their shunting yards, but in practical scenarios, only heuristics can solve TUSS instances (as was shown by Geiger et al. \cite{geiger2018}). Heuristic solutions are often sub-optimal and give little information about what optimal solutions would look like and what could still be achieved using existing shunting infrastructure. To gain more insight into the capacity of shunting yards and the feasibility of problem instances, Mulderij et al. \cite{mulderij2020} propose the multi-agent pathfinding (MAPF) problem \cite{stern2019} extended with a matching subproblem (hereafter MAPFM) as a suitable relaxation that can be used to compute upper-bounds for TUSS instances by solving the corresponding MAPFM instances optimally.

	Comparatively, little research has been done on MAPFM, and the Conflict-Based Min-Cost-Flow algorithm by Ma and Koenig \cite{ma2016} is one of the few existing approaches to optimally solve this more general problem along with the somewhat similar one described in \cite{henkel2019}. Ma and Koenig's method builds upon the conflict-based search method and in particular on the Meta-Agent variation thereof also discussed in \cite{sharon2015} as a high-level search framework, while exploiting the connection between MAPF problems and max-flow problems (as discussed in \cite{yu2013}) in the low-level search. It begs the question: can alternative MAPF algorithms such as Increasing Cost Tree Search (ICTS) \cite{sharon2011} or M* \cite{wagner2011} also serve as the basis for a MAPFM solver?

	This work is the outcome of a search for an efficient algorithm for optimally solving MAPFM derived from ICTS as an alternative to the approach outlined above\footnote{To the knowledge of the author, to date, no such algorithm is described in the literature}. ICTS, in short, is a two-level approach to MAPF with a top-level breadth-first traversal of an Increasing Cost Tree (ICT) representing combinations of per-agent path costs and a bottom-level evaluation of ICT nodes using multi-valued decision diagrams (MDDS). These MDDS represent per agent all possible paths to the goal of a set length. In the context of a research project in which multiple such MAPF algorithms were taken as starting points for MAPFM algorithms within a group, the novel ICTS-based algorithm is compared to these algorithms on a set of benchmarks similar to those described in \cite{stern2019}. In addition, it is analysed in terms of complexity and compared to both an enumerative approach that solves all matchings using regular ICTS and the algorithm described in \cite{ma2016}. Lastly, completeness of the algorithm is considered as this is a key property in the context of MAPFM as a TUSS relaxation. Regarding the relevance of this work, it should be noted that besides TUSS, there are other applications such as planning warehouse robots \cite{wurman2007} that also could benefit from novel algorithms for MAPFM.

	\section{Multi-agent pathfinding with matching} % Problem description
	MAPFM is the problem of finding agent-goal matchings that induce a non-conflicting routing of agents to their goals. Each matching corresponds to a MAPF instance, so each feasible matching has an optimal solution, i.e. a solution with a minimal Sum of Individual Costs (SIC). In approaching MAPFM as a TUSS relaxation to find upper-bounds for instances, optimal MAPFM solutions are the most interesting as they can characterise such upper-bounds. Furthermore, defining optimality as minimal SIC corresponds more closely to TUSS objective functions such as personnel cost than the other common MAPF objective, minimal makespan. Optimally solving MAPFM is equivalent to finding a feasible matching such that there is no other matching for which the corresponding MAPF instance has a lower optimal SIC. This motivates the joint nature of MAPFM: some matchings might be infeasible and others may have sub-optimal SIC costs, so especially in larger instances where there may be thousands of matchings, it is preferable to integrate the generation of matchings with the search for non-conflicting path-combinations to avoid having to exhaustively process all matchings. In this work, both integrated approaches and exhaustive methods using ICTS are considered.

	\subsection{Formulation}
	Formally, a MAPFM instance can be described as follows. Let $G = (V,E)$ be an undirected connected graph, with each $v\in V$ representing an obstacle-free tile on a 4-grid and each $e = (u,v)\in E$ representing a legal uniform-cost move between two such tiles. Let there be $K$ teams $t_1,\ldots, t_K$, where $t_i$ consists of $k_i$ agents with start positions $s_1^i,\ldots,s_{k_i}^i$ together with an equal number of team goals $g_1^i,\ldots,g_{k_i}^i$.

	Within each team $t_i$, agent $a_j^i$ are to be \textit{matched} to a unique goal $g_k^i$, so that exactly one agent is assigned to each goal within team $i$; once all agents are matched to goals, for each agent $a_j^i$ matched to $g_k^i$, a path from $s_j^i$ to $g_k^i$ is to be found such that all agents paths taken together are non-conflicting and therefore make up a solution \cite{ma2016}. In this work, non-conflicting means, in the conflict-terminology of \cite{stern2019}, that there are no vertex conflicts (and thus no edge conflicts) and no swapping conflicts. Practically, this means that no two agents may be at $v\in V$ at the same timestep and that if agents $i,j$ are at locations $u,v$ respectively at timestep $t$, they may not be at locations $v,u$ at timestep $t+1$.

	\section{Extending ICTS with matching} % Your contribution
	Increasing Cost Tree Search with Matching (ICTS-m) generalizes ICTS to optimally solve MAPFM instances. Few changes to ICTS were necessary to accomodate for matching in this increasing cost tree scheme. Specifically, the ICT root generation and the MDD generation had to be changed. In regular ICTS, given a MAPF instance with agents $a_1,\ldots,a_k$ with goals $g_1,\ldots,g_k$ starting at locations $s_1,\ldots,s_k$, the root is given by the cost $c_i$ for each agent $a_i$ of the shortest path from $s_i$ to $g_i$, i.e. without taking conflicts into account. Any optimal pathfinding algorithm can be used for this. The root of the ICT then is the vector $(c_1,c_2,\ldots,c_k)$ with $k$ children $(c_1 + 1, c_2,\ldots,c_k), (c_1, c_2 + 1,\ldots,c_k),\ldots,(c_1, c_2,\ldots,c_k + 1)$ so that each component of the parent is incremented by 1 in exactly one child. This child-generation rule holds for all ICT nodes.

	As described in \cite{sharon2011}, given $a_i$, generating all paths $(s_i,g_i)$ of a certain target cost $c_i$ can be accomplished by a breadth-first search on the $c_i$-steps time-expanded graph. Storing the resulting paths in an MDD, which has at most $|V|$ nodes at each timestep or depth, avoids the cost of storing all paths explicitly and also facilitates the efficient search of the $k$-agent space of path combinations.
	\subsection{Root generation for matching}
	Given a team $t_i$ of size $k_i$, for any agent $a_j^i$ there are $k_i$ costs $c_j(1),c_j(2),\ldots,c_j(k_i)$ for shortest paths from $s_j^i$ to each goal $g_k^i$. Making no assumptions about what goals other agents in $t_i$ are assigned to, it could be that in the optimal solution $a_j^i$ is assigned to a goal $g_{m}^i$ with cost $c_j(m)$ such that $c_j(m) = \min_{n\in\{0,\ldots,k_i\}} c_j(n)$. Therefore, to guarantee optimality, this minimal cost to any matching goal has to take the place of the shortest path cost from the original ICTS root. The resulting root is equivalent in the case that all teams are singleton. Intuitively, the sum of individual costs of this root is highly optimistic. Often, agents will not move to their closest matching goal in an optimal solution, even without considering conflicts.

	\subsection{MDD generation for matching}
	In MDD generation, each agent also has to consider not one goal but all matching goals. By following a BFS-based approach, with all team goals being goals in the search and in turn endpoints from which, using the time-expanded graph, the MDD was generated that is used in the low-level search, agent MDDs can be made to represent all possible paths of a set length to any matching goal. Once again, if all teams are singleton, the MDDs generated using this goal-set approach are the same as in regular ICTS.

	\subsection{Properties of ICTS-m}
	<insert proof>
	\printbibliography
